---
title: "Single file MARLOWE with PEAKS on Sample: Fish-DeNovo "
author: "Daniel Vogel"
date: "`r Sys.Date()`"
output:
  html_notebook:
    df_print: paged
    toc: yes
    toc_depth: 2
    toc_float: yes
---

```{r setup, include=FALSE}
library(MakeSearchSim)
library(CandidateSearchDatabase)
library(OrgIDPipeline)
library(tidyverse)


conn_list <- list(
  "dbname"= "candidate",
  "host" = "localhost",
  "port" = 3306,
  "user" = "msdba",
  "password" = "MassSpec2021!"
)


knitr::opts_chunk$set(echo = TRUE)
```





# Load/Generate overlap matrices and taxonomic groups

Load the RData files that contain the organism_metadata dataframe and the
overlap_mat_for_scoring dataframe.
```{r load-matrices-and-taxon-groups}
#must create variable named organism_metadata
metadata_file <-file.choose() 
load(metadata_file)
#must create variable named overlap_mat_for_scoring
scoring_overlap_file <- file.choose()
load(scoring_overlap_file)
```


Un-comment (Select lines then Ctrl-Shift-C) and run the below chunk if you don't
have the necessary matrices already saved. They can be saved and re-used until
new organisms are added to the database. Save the overlap_mat_for_scoring and
organism_metadata dataframes.
```{r generate-matrices-and-taxon-groups}
# organism_metadata <- CandidateSearchDatabase::organism_info(conn_list)
# 
# # Use this matrix for taxonomic groups
# overlap_mat <- CandidateSearchDatabase::overlap_coefficients(conn_list) %>%
#   column_to_rownames("kegg_id") %>%
#   replace(is.na(.), 1.)
# 
# # Use this matrix for scoring
# overlap_mat_for_scoring <- 
#   CandidateSearchDatabase::overlap_coefficients_strong(conn_list) %>%
#   column_to_rownames("kegg_id") %>%
#   replace(is.na(.), 1.)
# 
# grouping <- CandidateSearchDatabase::leader_similarity_cluster(
#   as.matrix(overlap_mat),
#   minimum_overlap_coefficent = .5
# )
# 
# organism_metadata <- organism_metadata %>%
#   left_join(
#     grouping,
#     by=c("kegg_id" = "org_id")
#   )
# rm(grouping)
```

# Generate Tags from PEAKS

PEAKS generates a csv which is named `de novo peptides.csv` by default. 
**The below chunk expects it to only contain results for one source file.** 
If it contains results for multiple source files see the 'Batch PEAKS MARLOWE
run' R markdown template.

A `.RData` file containing the tags will also be saved
in the designated `output_dir` directory.
```{r peaks-single-file}

peaks_output_path <- "PEAKS/denovo.Salvelinus_namayacush.csv"

#folder to hold denovo tag dataframes
tag_output_dir <- file.path(
      normalizePath(results_folder, winslash = "/"),
      "denovo_tags",
      fsep = "/"
    )

peaks_output <- read.csv(peaks_output_path,
                         header = TRUE, 
                         stringsAsFactors = FALSE)

#filter for ALC and peptide length
#standard filters are for a total peptide length >= 6
#and an ALC >= 50
peaks_output <- peaks_output %>%
    dplyr::filter(.data$length >= 6 &
                    .data$ALC.... >= 50)

tags_mass_factor_df <- generate_PEAKS_tags(all_peaks_denovo = peaks_output,
                                          output_dir = tag_output_dir,
                                          LC_threshold = 80, 
                                          prolineBlock = TRUE, 
                                          min_length = 5, 
                                          ppm_error = 15, 
                                          return_tags = TRUE)

```


# Uploading tags to database

`sample_id` must be a unique string, if a sample with the same sample_id already
exists in the database it will be overwritten.

**add sample_id string**
```{r upload-sample}

sample_id_string <- tools::file_path_sans_ext( names(table(peaks_output$Source.File))[1] )
print( paste0("sample_id_string: ", sample_id_string)) 

CandidateSearchDatabase::upload_sample(
  conn_list,
  sample_id = sample_id_string,
  sample = tags_mass_factor_df
  )
```

# Downloading and filtering database matches

If `delete_sample_on_completion = TRUE` (default) then the tags for each sample
will be deleted from the database once the results are downloaded. This prevents
queries from slowing down as the tables that store the sample data get bigger.

**add sample_id string** if restarting from here
```{r download-filter-db-matches}

#if restarting from here re-specify sample_id
# sample_id_string <- 

#output_db_dir <- choose.dir()
#folder to write raw tag-peptide-protein matches and count matrices
output_db_dir <- file.path(
      normalizePath(results_folder, winslash = "/"),
      "raw_matches_strong_counts",
      fsep = "/"
    )

#folder to write filtered count matrices
filt_counts_dir <- file.path(
      normalizePath(results_folder, winslash = "/"),
      "filtered_counts",
      fsep = "/"
    )
#filt_counts_dir <- choose.dir()

raw_hits_paths <- 
  CandidateSearchDatabase::save_database_hits(sample_id = sample_id_string, 
                                     conn_list = conn_list, 
                                     output_dir = output_db_dir,
                                     delete_sample_on_completion = TRUE)

#filtering database hits
filt_counts_path <- filter_counts(strong_count_file = raw_hits_paths["strong_counts"],
                                  pep_match_file = raw_hits_paths["peptide_matches"],
                                  min_pep_per_prot = 2, 
                                  output_dir = filt_counts_dir)

```


# Calculating Taxonomic Scores

If `fasta_for_filter_path` is not NULL all tags that are found within the
specified fasta files are removed.

`min_group_hits` is used to filter taxonomic groups before scoring. If a
taxonomic group has total hits < `min_group_hits` it is removed prior to
calculating pos_scores and org_scores. If there is only one taxonomic group that
passes the `min_group_hits` filter then  pos_scores and org_scores won't be
calculated and those columns won't exist in the output.

```{r calculate-scores}
fasta_for_filter_path <- "crap.fasta"  

#folder to save taxonomic scores
taxonomic_scores_output_folder <- file.path(
      normalizePath(results_folder, winslash = "/"),
      "taxonomic_scores",
      fsep = "/"
    )

#taxonomic_scores_output_folder <- choose.dir()

score_file <- run_organism_scoring(
    filt_counts_path,
    fasta_for_filter = fasta_for_filter_path,
    organism_metadata_df = organism_metadata,
    scoring_overlap_mat = overlap_mat_for_scoring,
    output_dir = taxonomic_scores_output_folder,
    Pthrsh = 0.4,
    min_total_tags = 8,
    min_total_strong_tags = 8,
    min_organism_strong_tags = 1,
    min_group_hits = 1,
    export_matrix = TRUE
  )
```


# Generating Heatmaps

## Organism level heatmap
Set `min_pos_score` and `min_org_hits` parameters to prevent too many organisms
from being included in the plot. The values of these parameters will be included
in the output filename.

Use the `overlap_mat` generated in Chunk 3: generate-matrices-and-taxon-groups
to visualize the overlap coefficients based on all peptides. Use the
`overlap_mat_for_scoring` to visualize the overlap coefficients based on strong
peptides only.


```{r}
#load overlap mat
# load()

omap <- organism_heatmap(score_file_path = score_file,
                      overlap_matrix = overlap_mat,
                      output_dir = "heatmaps",
                      output_type = "png",
                      min_org_hits = ,
                      min_pos_score = ,
                      width = 10,
                      height = 7,
                      units = "in")


plot(omap)
```

## Group level heatmap
The `_reduced_matrix.csv` file will be in the `taxonomic_scores_output_folder`
if `export_matrix = TRUE` in the `run_organism_scoring()` call above.

Set the `min_pos_score` parameter to prevent too many organisms
from being included in the plot. The value of this parameter will be included
in the output filename.

```{r}
reduced_matrix_csv <- gsub("organism_scores.csv", 
                           "reduced_matrix.csv",
                           score_file, 
                           fixed = TRUE)

gmap <- group_heatmap(score_file_path = score_file,
                      reduced_matrix_path = reduced_matrix_csv,
                      output_dir = "heatmaps",
                      output_type = "png",
                      min_pos_score = 1 )

plot(gmap)
```




# Reproducibility Information
```{r colophon}
sessionInfo()
```

